---
title: "Portfolio Managment: Homework 1"
author: "Gal Skarishevsky, Nathan Matare, Lior, Sahaf"
date: "March 30, 2017"
output: pdf_document
urlcolor: blue
graphics: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE,
                      fig.width = 6, fig.height = 4.5, fig.align = "right")
```

```{r Load Config, include = FALSE, echo = FALSE, results = 'hide'}

options("width" = 250)
options(scipen  = 999)
options(digits  = 003)

library(xts); library(zoo); library(e1071); library(ggplot2); library(knitr)

set.seed(666) # the devils seed

username <- Sys.info()[["user"]]
dir <- paste("/home/", username, "/Documents/Education/Chicago_Booth/Classes/35120_Portfolio_Management/portfolio_management/homework1", sep = "")
setwd(dir)
```

# 1. Load Data

We write the function '\texttt{getReturns}' that loads the data into our R environment.

```{r load data, include = TRUE, echo = TRUE}

getReturns <- function(name){
returns <- read.csv(name, skip = 4, sep = "\t", header = FALSE)
				
if(name == 'returns_annual.txt') 
	returns <- as.xts(returns[ ,-1], 
					      order.by = as.Date(as.character(returns[,1]), format = "%Y"))
else 
	returns <- as.xts(returns[ ,-1], 
			          order.by = as.Date(as.character(returns[,1]), format = "%Y%m%d"))
				
colnames(returns) <- c("SP500_rtrn", "BOND_rtrn")
return(returns)
}

daily <- getReturns('returns_daily.txt')
monthly <- getReturns('returns_monthly.txt')
annually <- getReturns('returns_annual.txt')

```

# 2 and 3. Estimate basic statistics

We write the function '\texttt{getStats}' that computes the means, variances, standard deviations, covariances, correlations, skewness, and kurtosis of each time series at all three frequences.

```{r getStats1, include = TRUE, echo = FALSE}
getStats <- function(DF){
				stock <- DF$SP500_rtrn
				bond <- DF$BOND_rtrn

				stats <- data.frame(

							stock = c(	mean = mean(stock), var = var(stock), 
										sd = sd(stock), skew = skewness(stock), 
										kurt = kurtosis(stock)
									),

							bond  = c(	mean = mean(bond), var = var(bond), 
										sd = sd(bond), skew = skewness(bond), 
										kurt = kurtosis(bond)
									)

				)

				shape <- c(cov = cov(stock, bond), cor = cor(stock, bond))
				return(list(stats = stats, shape = shape))
}
```

```{r getStats2, include = TRUE, echo = FALSE}

kable(getStats(daily)$stats, digits = 4, caption = "Daily")
kable(t(getStats(daily)$shape), digits = 6)

kable(getStats(monthly)$stats, digits = 4)
kable(t(getStats(monthly)$shape), digits = 6, caption = "Monthly")

kable(getStats(annually)$stats, digits = 4)
kable(t(getStats(annually)$shape), digits = 6, caption = "Annually")

```

\newpage
# 4. Histograms

None of frequencies (daily, monthly or annually) are normally distributed. We observe greater non-normality the larger the time frequencies become. See the below histograms. The S&P500 returns are graphed in blue while bonds are graphed in red.

```{r do Hist, include = TRUE, echo = FALSE, fig.height=4.5, fig.width=8.5, fig.align='center'}

ggplot(data = daily) +
	geom_histogram(aes(SP500_rtrn), fill = "blue", color = "green", alpha = 0.5) +
	geom_histogram(aes(BOND_rtrn), fill = "red", color = "red", alpha = 0.5)

ggplot(data = monthly) +
	geom_histogram(aes(SP500_rtrn), fill = "blue", color = "green", alpha = 0.5) +
	geom_histogram(aes(BOND_rtrn), fill = "red", color = "red", alpha = 0.5)

ggplot(data = annually) +
	geom_histogram(aes(SP500_rtrn), fill = "blue", color = "green", alpha = 0.5) +
	geom_histogram(aes(BOND_rtrn), fill = "red", color = "red", alpha = 0.5)

```

# 5. Compute CIs

We write the function '\texttt{getADVStats}' that computes the upper and lower confidence intervals for a period of one, and a rolling arithmetic period of thirty. 

```{r getCIs, include = TRUE, echo = TRUE}

getADVStats <- function(DF){
stock <- DF$SP500_rtrn
bond <- DF$BOND_rtrn

error <- qnorm(0.975) * sd(stock) / sqrt(NROW(stock))
CI_1 <- c(lower = mean(stock) - error, upper = mean(stock) + error)

stock_30 <- na.omit(rollapply(stock, width = 30, mean))
error_30 <- qnorm(0.975) * sd(stock_30) / sqrt(NROW(stock_30))
CI_30 <- c(lower = mean(stock_30) - error_30, upper = mean(stock_30) + error_30)
return(data.frame(period_1 = CI_1, period_30 = CI_30))
}
```


```{r getCIs2, include = TRUE, echo = FALSE}

kable(getADVStats(daily), digits = 6, caption = "Daily")
kable(getADVStats(monthly), digits = 6, caption = "Monthly")
kable(getADVStats(annually), digits = 6, caption = "Annually")

```

# 6. Compute Absolute Shortfalls

We write the function '\texttt{getABSshortfall}' that computes the probability that the return over the next period will be lower than a specified threshold parameters, k. 

```{r absShortfalls, include = TRUE, echo = TRUE}

getABSshortfall <- function(DF){
	stock <- DF$SP500_rtrn
	bond <- DF$BOND_rtrn

	ks <- c(-0.20, -0.10, 0, 0.10, 0.20)
	out <- matrix(NA, 5, 2, dimnames = list(ks, c("stock", "bond")))

	for(k in 1:NROW(out)){
			p <- as.numeric(rownames(out)[k])
			out[k, 1] <- pnorm(p, mean = mean(stock), sd = sd(stock))
			out[k, 2] <- pnorm(p, mean = mean(bond), sd = sd(bond))
	}

	return(out)
}
```

```{r absShortfalls2, include = TRUE, echo = FALSE}

kable(getABSshortfall(daily), digits = 6, caption = "Daily")
kable(getABSshortfall(monthly), digits = 6, caption = "Monthly")
kable(getABSshortfall(annually), digits = 6, caption = "Annually")

```

# 7. Stock Return Probability

We write the function '\texttt{getStockBondProb}' that computes the probability that the stock return is lower than the bond return for a given frequency. 

```{r getStockBondProb, include = TRUE, echo = TRUE}

getStockBondProb <- function(DF){
				stock <- DF$SP500_rtrn
				bond <- DF$BOND_rtrn
        
				# number of times stock return is lower than bond return
				num_times <- length(which(stock < bond)) 
				# number of times event happened over number of observations
				prob <- num_times / NROW(DF) 
				return(prob)
}
```

```{r getStockBondProb2, include = TRUE, echo = FALSE}

c(Daily = getStockBondProb(daily))
c(Monthly = getStockBondProb(monthly))
c(Annually = getStockBondProb(annually))

```
