---
title: 'Portfolio Managment: Homework 8'
author: "Gal Skarishevsky, Nathan Matare, Brian Thompson, Lior Sahaf"
date: "May 18, 2017"
output: pdf_document
graphics: yes
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE,
                      fig.width = 6, fig.height = 4.5, fig.align = "right")
```

```{r Load Config, include = FALSE, echo = FALSE, results = 'hide'}
options("width" = 250)
options(scipen  = 999)
options(digits  = 003)

library(xts); library(zoo); library(e1071); library(R.matlab)
library(ggplot2); library(knitr); library(gridExtra); library(Hmisc)
library(reshape2); library(foreach); library(doMC); library(data.table)

set.seed(666) # the devils seed

username 	<- Sys.info()[["user"]]
dir 		<- paste("/home/", username, "/Documents/Education/Chicago_Booth/Classes/35120_Portfolio_Management/portfolio_management/mod8", sep = ""); setwd(dir)

rets  <- t(readMat("rets_hwk7.mat")$rets)[-1, ] # -99 are NA values
rets[rets == -99] <- NA
colnames(rets) <- c("Date", paste("Fund", rep(1:(dim(rets)[2] - 1))))
rets <- as.data.table(rets)

flows <- t(readMat("flows_hwk7.mat")$flows)[-1, ] # first one is fund identifier
flows[flows == -99] <- NA
colnames(flows) <- c("Date", paste("Fund", rep(1:(dim(flows)[2] - 1))))
flows <- as.data.table(flows)

```

# B.1. Estimating E and V by the sample estimates.

stuff

# B.2. Estimating the Performance-Flow Relation.

We write the function '\texttt{EVALperformancetoflow}' that computes our desired question answers:

```{r function_setup2, include = TRUE, echo = TRUE}

EVAL_performance_to_flow <- function(year, graph = FALSE, ...){
	data <- transpose(rbind(rets[Date == year], flows[Date == year + 1]))
	colnames(data) <- c("returns", "inflows")
	data <- na.omit(data[-1, ]) # remove year and remove NAs

	data <- data[order(returns, decreasing = TRUE)]
	data[ ,group_interval := cut_number(returns, n = 10)]
	data[, decile := as.integer(group_interval)]
	data[ ,avg_returns := mean(returns), by = group_interval]
	data[ ,avg_flows := mean(inflows), by = group_interval]

	data_per_decile <- data[!duplicated(avg_returns)]
	linear <- lm(avg_flows ~ avg_returns + I(avg_returns ^ 2), data = data_per_decile)

	data_per_decile[ ,hat_avg_flows := linear$fitted][ ,c("returns", "inflows") := NULL]

	estimates <- rbind(
					alpha_hat = summary(linear)$coefficients[,'Estimate'][1], 
					beta_hat = summary(linear)$coefficients[,'Estimate'][2], 
					charlie_hat = summary(linear)$coefficients[,'Estimate'][3],

					alpha_se  = summary(linear)$coefficients[,'Std. Error'][1], 
					beta_se  = summary(linear)$coefficients[,'Std. Error'][2], 
					charlie_se  = summary(linear)$coefficients[,'Std. Error'][3]
	); colnames(estimates) <- year

	data <- merge(data, data_per_decile)

	if(graph){

		p <- ggplot(data = data, aes(decile, group = 1)) + 
				geom_point(aes(y = avg_flows), color = "black") + 
				geom_line(aes(y = avg_flows)) +
				geom_line(aes(y = hat_avg_flows), color = "darkgrey", linetype = "dotted", size = 1.3) +
				ggtitle(paste("Performance(", year, ") to Fund Flows(", year + 1, ")", sep = "")) +
				labs(x = "Performance Deciles", y = "Average Fund Flow (t+1)") 
		print(p)
	}

	return(list(data = data_per_decile, estimates = estimates))
}

```

## a) i - iv. 
 
We evaluate the performance-fund for 2001 and observe a convex pattern. As expected, the better performing the fund in the previous year, the more new funds flow into it. Further, our estimates (fitted_values) are highly correlated with the real values. This linear model explains much of the variation (in-sample, of course). All three of the coefficients appear to be highly significant. 

```{r 2i, include = TRUE, echo = FALSE}
out <- EVAL_performance_to_flow(year = 2001, graph = TRUE)
```

```{r 2ii, include = TRUE, echo = FALSE}
kable(out$data, digits = 4, caption = "Average: Returns, Flows, Fitted Values by Decile")
```

```{r 2iii, include = TRUE, echo = FALSE}
kable(t(out$estimates), digits = 4, caption = "Coefficients and Standard Errors")
```

## b) i.

We write the function '\texttt{getStatSignif}' and compute our Fama-MacBeth estimates:

```{r function_setup3, include = TRUE, echo = TRUE}
years <- as.numeric(unlist(rets[,'Date'][-11])) # remove 2002
performance <- lapply(years, EVAL_performance_to_flow)

getStatSignif <- function(coef, ...){
	FM_alpha_hat <- do.call(sum, lapply(performance, function(x) 
	                    x$estimates[paste(coef, "_hat", sep = ""),])) / 10
	FM_alpha_se  <- sd(unlist(lapply(performance, function(x) 
	                    x$estimates[paste(coef, "_se", sep = ""),]))) / sqrt(10)
	summary_stat <- cbind(FM_alpha_hat, FM_alpha_se)
	colnames(summary_stat) <- NULL; rownames(summary_stat) <- coef
	return(summary_stat)
}

```

Confirming our results from above, the Fama-Macbeth approach is significant for all three coefficients, implying that the convexity pattern(charlie) is statistically signficant as well.

```{r i, include = TRUE, echo = FALSE}
coefs <- c("alpha", "beta", "charlie")
coef_stats <- t(sapply(coefs, getStatSignif))
colnames(coef_stats) <- c("Estimate", "Standard Error")
kable(coef_stats, digits = 4, caption = "Fama-MacBeth Estimates")
```

## b) ii

Next, we average the performance-flow per decile across years. We, again, observe a convex pattern:

```{r iasdf, include = FALSE, echo = FALSE}
total_summary <- do.call(rbind, lapply(performance, function(x) x$data))
total_summary[ ,total_avg_flow := mean(avg_flows), by = decile]
summary_data <- total_summary[!duplicated(decile)]
```

```{r ii, include = TRUE, echo = FALSE}
ggplot(data = summary_data, aes(decile, group = 1)) + 
	geom_point(aes(y = avg_flows), color = "black") + 
	geom_line(aes(y = avg_flows)) +
	ggtitle("Total Average Performance to Flow") +
	labs(x = "Performance Deciles", y = "Average Fund Flow (t+1)") 
```

## c) i. 

$\mathbb{E}$(A) = 1(0.01) + 0(0) = 0.01  
$\mathbb{E}$(B) = -0.25(0.60) + 0.25(0.40) = -0.05

Thus, choose A.

## c) ii. 

Given,

Initial investment size = 100M  
$\hat{a}$ = 0.04  
$\hat{b}$ = 1.56  
$\hat{c}$ = 4.44

Scenario A:  
$\hat{F_A}$ = 0.04 + 1.56 * 0.01 + 4.44(0.01^2) = 0.05  
Year-end investment size = 105.6  
$\mathbb{E_A}$(A) = 1.05M

Scenario B:  
$\hat{F_B}$ = 0.04 + 1.56 * 0.25 + 4.44(0.25^2) = 0.7075  
Year-end investment size = 170.75  
Fee(good) = 1.70M  

$\hat{F_B}$ = 0.04 + 1.56 * -0.25 + 4.44(-0.25^2) = -0.0725  
Year-end investment size = 92.75
Fee(bad) = 9.275M   

$\mathbb{E}$(B) = 0.40(1.70) + 0.60(0.92) = 1.23

Thus, choose B.

## c) iii. 

Due to the quadractic term and incentive(1% fee after inflows) structure, the potential upside for investment B, even after factoring in the investment's riskyness (-0.25), is greater (1.23 > 1.05). Thus, a short term looking money manger would have greater incentives to take risky bets given their expected payoff. 

## d) 

New money chases hot money.  

There holds a strong quadratic relationship between a funds past return and its future inflows. Although it may be that these well-performing funds continue to perform well into the future (more analysis needed), given our class discussion, notes, and referenced papers, this seems unlikely. And, noting that incentives are tied not only to raw return performance but also fund inflows, it appears that money managers stand to gain, in expectation, significant upside on risky investments.  
